---
title: "NYPD Shooting Incident"
author: ''
date: "23/10/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import Library

```{r library_imp, warning = FALSE, message = FALSE}
  library(tidyverse)
  library(lubridate)
  library(scales) 
```

## Import Data

Importing the NYPD Shooting Incident Data form [Catalog Data](https://catalog.data.gov/dataset)


```{r import_data, echo=FALSE, include=FALSE}
  
  url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
  nypd_data = read_csv(url)
  nypd_data
```

### Import Summary

```{r import_summary}
  summary(nypd_data)
```

---

## Tidy

Identified the following variables necessary for my analysis on nypd_data.

- PERP_SEX 
- VIC_SEX
- BORO
- OCCUR_TIME
- OCCUR_DATE
- Longitude

Tidying the data by removing all other variables, and transforming the OCCUR_DATE from character format to Date Object.

```{r tidy_data, echo=FALSE, warning=FALSE}
nypd_data <- nypd_data %>% select(-c(INCIDENT_KEY,PRECINCT,JURISDICTION_CODE,LOCATION_DESC,PERP_RACE,VIC_AGE_GROUP,VIC_RACE,X_COORD_CD,Y_COORD_CD,Lon_Lat,PERP_AGE_GROUP,Latitude,STATISTICAL_MURDER_FLAG)) %>% 
  mutate(OCCUR_DATE = mdy(OCCUR_DATE))
# Changing date type to an date object

```

### Tidy Summary 

```{r tidy_summary}
  summary(nypd_data)
```

---

## Missing Data
In the following analysis, the missing data (NA/NR/U) has been ignored.

---

## Visualization

### Sex based comparison of Perpatrators 

```{r bar_plot_gend_perp}
x <- c("Male","Female")
y <- c(nrow(nypd_data %>% filter(PERP_SEX == 'M')),
       nrow(nypd_data %>% filter(PERP_SEX == 'F')))
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('cadetblue1','pink')
pie(y,label = perc,main = "Perpatrators", col = colors)
legend("bottomright", x, fill = colors)
```

### Sex based comparison of Victims 

```{r bar_plot_gend_vic}
x <- c("Male","Female")
y <- c(nrow(nypd_data %>% filter(VIC_SEX == 'M')),
       nrow(nypd_data %>% filter(VIC_SEX == 'F')))
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('dodgerblue1','gold')
pie(y,label = perc,main = "Victims", col = colors)
legend("bottomright", x, fill = colors)
```

### Sex based Victims ratio of Male Perpatrators

```{r pm_to_vic}
pmd <- nypd_data %>% filter(PERP_SEX == 'M') 
pmd_f_c <- nrow(pmd %>% filter(VIC_SEX == "F"))
pmd_m_c <- nrow(pmd %>% filter(VIC_SEX == "M"))
x <- c("Male","Female")
y <- c(pmd_m_c,pmd_f_c)
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('dodgerblue1','gold')
pie(y,label = perc,main = "Victims of Male Perpatrators", col = colors)
legend("bottomright", x, fill = colors)
```

### Sex based Victims ratio of Female Perpatrators

```{r pf_to_vic}
pmd <- nypd_data %>% filter(PERP_SEX == 'F') 
pfd_f_c <- nrow(pmd %>% filter(VIC_SEX == "F"))
pfd_m_c <- nrow(pmd %>% filter(VIC_SEX == "M"))
x <- c("Male","Female")
y <- c(pfd_m_c,pfd_f_c)
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('dodgerblue1','gold')
pie(y,label = perc,main = "Victims of Female Perpatrators", col = colors)
legend("bottomright", x, fill = colors)
```

### Sex based Perpatrators ratio of Male Victims

```{r vm_to_p}
pmd <- nypd_data %>% filter(VIC_SEX == 'M') 
pfd_f_c <- nrow(pmd %>% filter(PERP_SEX == "F"))
pfd_m_c <- nrow(pmd %>% filter(PERP_SEX == "M"))
x <- c("Male","Female")
y <- c(pfd_m_c,pfd_f_c)
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('cadetblue1','pink')
pie(y,label = perc,main = "Perpatrators of Male Victims", col = colors)
legend("bottomright", x, fill = colors)
```

### Sex based Perpatrators ratio of Female Victims

```{r vf_to_p}
pmd <- nypd_data %>% filter(VIC_SEX == 'F') 
pfd_f_c <- nrow(pmd %>% filter(PERP_SEX == "F"))
pfd_m_c <- nrow(pmd %>% filter(PERP_SEX == "M"))
x <- c("Male","Female")
y <- c(pfd_m_c,pfd_f_c)
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('cadetblue1','pink')
pie(y,label = perc,main = "Perpatrators of Female Victims", col = colors)
legend("bottomright", x, fill = colors)
```

### Date-Time Scatter plot 

```{r tim_plot}
  tt_data <- nypd_data %>% filter(OCCUR_DATE > '2018-12-31' & OCCUR_DATE < '2020-1-1')  
  tp <- ggplot(tt_data, aes(OCCUR_DATE, OCCUR_TIME)) +
           geom_point(na.rm=TRUE, color="blue", size=1) + 
           ggtitle("NYPD Shooting Data of 2019") +
           xlab("Date") + ylab("Time") 
  tpb <- tp + (scale_x_date(breaks = date_breaks("1 month"),labels=date_format("%b %m")))
  tt_data1 <- nypd_data %>% filter(OCCUR_DATE > '2019-12-31' & OCCUR_DATE < '2020-12-6')  
  tp1 <- ggplot(tt_data1, aes(OCCUR_DATE, OCCUR_TIME)) +
           geom_point(na.rm=TRUE, color="blue", size=1) + 
           ggtitle("NYPD Shooting Data of 2020") +
           xlab("Date") + ylab("Time")
  tpb1 <- tp1 + (scale_x_date(breaks = date_breaks("1 month"),labels=date_format("%b %m")))
  
  tpb
  tpb1
```

## Analysis of Date-Time Scatter Plot
The Shooting Incidents drop in the mid-morining(8 to 11 am). Shooting Incidents have increased in 2020.

### Borough wise Density Comparison

```{r boro_comp}
b_l <- unique(nypd_data$BORO)
y <- c(nrow(nypd_data %>% filter(BORO == b_l[1])),
       nrow(nypd_data %>% filter(BORO == b_l[2])),
       nrow(nypd_data %>% filter(BORO == b_l[3])),
       nrow(nypd_data %>% filter(BORO == b_l[4])),
       nrow(nypd_data %>% filter(BORO == b_l[5])))
perc <- paste0(round(100 * y/sum(y), 2), "%")
colors <- c('dodgerblue1','gold','lightcoral','lightcyan4','ivory1')
pie(y,label = perc,main = "Borough", col = colors)
legend("bottomright", b_l, fill = colors, cex = 0.7)
```

### Saftey Index of Boroughs
The index is in highest to lowest order.

1. Staten Island
2. Manhattan
3. Queens
4. Bronx
5. Brooklyn

### Model

A basic **Linear Model**, which predicts the **Longitude** of a crime on a specific date. This model uses the existing data to predict the outcome, which has been compared with the real outcomes.

```{r ln_mod}
m_data <- nypd_data %>% filter(OCCUR_DATE >= "2007-1-1" & OCCUR_DATE <= "2007-1-31")
mod <- lm(Longitude ~ OCCUR_DATE,data = m_data)
m_data_pred <- m_data %>% mutate(pred = predict(mod))
m_data_pred %>% ggplot() +
 ggtitle("January 2007") + 
 geom_point(aes(x = OCCUR_DATE,y = Longitude),color = "blue") +  
 geom_point(aes(x = OCCUR_DATE,y = pred),color = "red")   
```

## Conclusion

In this report the analyses, model and visualizations were on, how the Ratio of each provided Gender compared to each other in Perpetrators and Victims, and saw the period in which these incidents happen, and a Linear Model. For future work, many other analyses are possible, such as a **Demographics-based Visualization** or an **Age-Group-based Visualization**, etc.

## Session Info

```{r session_inf}
sessionInfo()
```
